---
title: "Scrape Novel Coronavirus (COVID-19) Cases, provided by JHU CSSE"
author: "Jawaid Hakim"
date: "`r Sys.Date()`"
output:
  
  html_document:
    
    toc: true
    toc_float: true
    number_sections: true
  pdf_document: 
    toc: true
    number_sections: true
boxlinks: true
urlcolor: blue
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load library

```{r}
library(tidyverse)
```

```{r}
source("./Consts.R")
source("./FilePaths.R")
```

# Generate list of data to be processed

# Setup parallel processing

```{r}
library(doParallel)
library(parallel)
```

Create data extraction functions to run on Spark cluster for processing COVID stats.

```{r par-table-funcs}
#
# Clean stats table
# Parameters:
#   statsTbl: stats table
#   targetDate: stats table date
#
# Returns:
#   Cleaned stats table
#
CleanStatsTable <- function(statsTbl, targetDate) {
    library(tidyverse)
    library(lubridate)
    
    # change column types
    cleanedTbl <- statsTbl %>%
        mutate(
            Confirmed = as.integer(Confirmed),
            Deaths = as.integer(Deaths),
            Recovered = as.integer(Recovered),
            Active = as.integer(Active),
            FIPS = as.numeric(FIPS),
            Incident_Rate = as.numeric(Incident_Rate),
            Case_Fatality_Ratio = as.numeric(Case_Fatality_Ratio),
            
            Last_Update = as_datetime(Last_Update)
        )
    
    target_year <- lubridate::year(targetDate)
    target_month <- lubridate::month(targetDate)
    target_day <- lubridate::day(targetDate)
    
    # add date/year/month/dat columns
    cleanedTbl <-  cleanedTbl %>%
        mutate(Date = targetDate, .after = Country_Region) %>%
        mutate(Year = target_year, .after = Date) %>%
        mutate(Month = target_month, .after = Year) %>%
        mutate(Day = target_day, .after = Month)
    
    return (cleanedTbl)
}

#
# Clean USA stats table
# Parameters:
#   statsTbl: stats table
#   targetDate: stats table date
#
# Returns:
#   Cleaned stats table
#
CleanStatsTable.USA <- function(statsTbl, targetDate) {
    library(tidyverse)
    library(lubridate)

    cleanedTbl <- CleanStatsTable(statsTbl = statsTbl, targetDate = targetDate)
    cleanedTbl <- cleanedTbl %>%
        mutate(
            Total_Test_Results = as.integer(Total_Test_Results),
            People_Hospitalized = as.integer(People_Hospitalized),
            UID = as.integer(UID),
            Testing_Rate = as.numeric(Testing_Rate),
            Hospitalization_Rate = as.numeric(Hospitalization_Rate)
        )
    return (cleanedTbl)
}

#
# Clean Global stats table
# Parameters:
#   statsTbl: stats table
#   targetDate: stats table date
#
# Returns:
#   Cleaned stats table
#
CleanStatsTable.Global <- function(statsTbl, targetDate) {
    library(tidyverse)
    
    cleanedTbl <- CleanStatsTable(statsTbl = stats_tbl, targetDate = targetDate)
    return (cleanedTbl)    
}

```


```{r par-scrape-funcs}

ScrapeStats <- function(url_index) {
    
    # need to load libraries on cluster nodes
    library(tidyverse)
    library(readr)

    source("./Consts.R", local = TRUE)

    # get URL and date from exported namespace
    targetUrl <- tbl_href$raw_href[url_index]
    targetDate <- tbl_href$date[url_index]

    stats_tbl <- readr::read_csv(file = targetUrl)
    
    return (stats_tbl)    
}

#
# Scrape Global stats table
# Parameters:
#   url_index: index of table to process
#
# Returns:
#   Scraped stats table
#
ScrapeStats.Global <- function(...) {
    
    library(tidyverse)
    source("./FilePaths.R", local = TRUE)

    # get index of URL to process
    url_index <- (...)
    targetDate <- tbl_href$date[url_index]
    
    stats_tbl <- ScrapeStats(url_index)
    
    cleanedTbl <- CleanStatsTable.Global(statsTbl = stats_tbl, targetDate = targetDate)
    
    path <- Output.GetDataFilePath(targetDate = targetDate, global = FALSE)
    write_csv(cleanedTbl, file = path, quote = 'needed')
    
    return (cleanedTbl)
}

#
# Scrape USA stats table
# Parameters:
#   url_index: index of table to process
#
# Returns:
#   Scraped stats table
#
ScrapeStats.USA <- function(...) {
    
    library(tidyverse)
    source("./FilePaths.R", local = TRUE)

    # get index of URL to process
    url_index <- (...)
    targetDate <- tbl_href$date[url_index]
    
    stats_tbl <- ScrapeStats(url_index)
    
    cleanedTbl <- CleanStatsTable.USA(statsTbl = stats_tbl, targetDate = targetDate)
    
    path <- Output.GetDataFilePath(targetDate = targetDate, global = FALSE)
    write_csv(cleanedTbl, file = path, quote = 'needed')
    
    return (url_index)
}

```

Source local cluster functions.

```{r souce-cluster-functions}
source("./LocalClusterFunctions.R")
```

Start cluster
```{r start-cluster}
cr <- Cluster.Start()
```


```{r load-hrefs}
href_path <- Output.GetHrefFilePath('href', global = FALSE)
tbl_href <- read_csv(href_path)
tbl_href
```

Export namespace to cluster nodes
```{r export-namespace}
# Function ScrapeStats.USA will be run on cluster nodes. It will access to below namespaces
Cluster.ExportNamespace(cr, list('ScrapeStats', 'ScrapeStats.USA', 'ScrapeStats.Global', 'CleanStatsTable', 'CleanStatsTable.USA', 'CleanStatsTable.Global', 'tbl_href'))
```

Generate sequence id for all URLs. Each work item - i.e. URL to be processed - will be accessed by it's index in `tbl_href` table.

```{r generate-url-id}

# Which URLs to process. For full set: start_href = 1, max_hrefs = NROW(tbl_href)
start_href <- 1
max_href <- 2

# For dev/testing purposes create a small sequence. For example, a sequence of 1:5 will process first 5 URLs
# For the final run use seq_along(1:NROW(tbl_href))
seq_id <- seq(from = start_href, length.out = max_href)
glimpse(seq_id)
```

Process URLs using cluster (map/reduce)
```{r par-process}
work_results <- Cluster.ParApply(cr, seq_id, ScrapeStats.USA)
work_results
```

Stop  cluster
```{r stop-cluster}
Cluster.Stop(cr)

```


```{r echo=FALSE aggregate-results}
#cum_results <- bind_rows(work_results)

# sort by descending date order
#cum_results <- cum_results %>% arrange(desc(Date))

#glimpse(cum_results)
```


```{r echo = FALSE save-processed-hrefs}
# save_set <- tbl_href[start_href:max_href, ]
# full_path <- Output.GetFullPath('href', global = FALSE)
# write_csv(save_set, full_path, quote = 'needed')
```

```{r echo=FALSE save-processed-data}
# full_path <- Output.GetFullPath('data', global = FALSE)
# write_csv(cum_results, full_path, quote = 'needed')
```

```{r echo=FALSE save-timestamp}
# full_path <- Output.GetFullPath('ts', global = FALSE)
# d <- format(date(), '%Y-%m-%d')
# write_file(s, full_path)

```

